---
title: "DATA 607, Assignment 07 -- XML and JSON in R"
author: "Kavya Beheraj"
date: "March 18, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(XML)
```

<br>

****

In this assignment, I recorded certain information for three different books, hand-made three files in HTML, XML, and JSON formats to store the information, and read the files into R to see how they compare.

These are the three books I chose:

* *The Dictator's Handbook*, by Bruce Bueno de Mesquita and Alastair Smith (2011)
* *King Leopold's Ghost*, by Adam Hochschild (1998)
* *The Complete Maus (Maus #1-2)*, by Art Spiegelman (1986)

Here are links to the raw files:

* [HTML file]("https://raw.githubusercontent.com/koffeeya/msds/master/DATA%20607%20Data%20Acquisition%20and%20Management/Assignments/Assignment%2007/books.html")

* [XML file]("https://raw.githubusercontent.com/koffeeya/msds/master/DATA%20607%20Data%20Acquisition%20and%20Management/Assignments/Assignment%2007/books.xml")

* [JSON file]("https://raw.githubusercontent.com/koffeeya/msds/master/DATA%20607%20Data%20Acquisition%20and%20Management/Assignments/Assignment%2007/books.json")

****

<br>

# HTML File

```{r}

u <- "https://raw.githubusercontent.com/koffeeya/msds/master/DATA%20607%20Data%20Acquisition%20and%20Management/Assignments/Assignment%2007/books.html"

html_df <- htmlParse(file = u)

print(html_df)

```


