---
title: 'Hands on Lab: Subsetting Datasets'
author: "Kavya Beheraj"
date: "February 3, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages("tidyverse", repos = "http://cran.us.r-project.org")

```

Your task is to work with the [Pittsburgh bridges dataset](https://archive.ics.uci.edu/ml/datasets/Pittsburgh+Bridges).

You should first study the data and its associated description (i.e. "data dictionary"). 

You should take the data, and create an R data frame with a subset of the columns in the dataset. You should provide relevant column names.

Your deliverable is the R Markdown code to perform these transformation tasks.  

***

##### **STEP 1: ** Read in the dataframe.

```{r}

bridges <- read.table("https://archive.ics.uci.edu/ml/machine-learning-databases/bridges/bridges.data.version2", sep=",")

```

***

##### **STEP 2: ** Rename the columns according to the dataset description.

```{r}

names(bridges) <- (c("IDENTIF", "RIVER", "LOCATION", "ERECTED", "PURPOSE", "LENGTH", "LANES", "CLEAR-G", "T-OR-D", "MATERIAL", "SPAN", "REL-L", "TYPE"))

```

***

##### **STEP 3: ** Create a subset of the data. In this case, I chose to only show bridges with a length of medium, and to limit the columns to all between "LOCATION" and "LENGTH".

```{r}

newdata <- subset(bridges, LENGTH == "MEDIUM", select=LOCATION:LENGTH)

print(newdata)

```

