---
title: "DATA 606 -- Project Proposal"
author: "Kavya Beheraj"
date: "April 12, 2018"
output:
  html_document:
    theme: journal
    highlight: haddock
    toc: true
    toc_depth: 4
    toc_float:
      collapsed: true
      smooth_scroll: true

---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

****

# 1 | Background

#### A | Libraries

```{r warning=FALSE, message=FALSE}
library(dplyr)
library(ggplot2)
library(psych)
```


<br>

# 2 | **Research Question**

### Is a character's sex predictive of their alignment in DC comics?


<br>

# 3 | Prepare the Dataset

#### A | Read in the data

```{r}
dc <- read.csv("https://raw.githubusercontent.com/fivethirtyeight/data/master/comic-characters/dc-wikia-data.csv", header = TRUE, sep = ",")
```

```{r}
head(dc, 3)
```

#### B | Pull out relevant variables


<br>

# 4 | About the Data

#### A | What are the cases, and how many are there?

Each case represents a character in the DC comics universe. There are 6,896 observations.

****

#### B | Describe the method of data collection.

The data was web-scraped by FiveThirtyEight. It was scraped from August 24 --- October 6, 2014. 

****

#### C | What type of study is this?

This is an observational study.

****

#### D | Data Source


<br>

# 5 | About the Variables

#### A | Response

The response variable is character alignment. It is categorical with the following levels: Good, Bad, Neutral, and Reformed Criminals.

****

#### B | Explanatory

The explanatory variable is character sex. It is a categorical variable with four levels: Female, Genderless, Male, and Transgender.


<br>

# 5 | Summary statistics

```{r}
summary(dc$ALIGN)
summary(dc$SEX)
```

