---
title: "DATA 606 - Chapter 1 Practice"
author: "Kavya Beheraj"
date: "February 6, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
install.packages(c('openintro','OIdata','ggplot2','psych','reshape2',
				   'knitr','markdown','shiny'), repos = "http://cran.us.r-project.org")
library('DATA606')
```

Practice - 1.7, 1.9, 1.23, 1.33, 1.55, 1.69

****

### **1.7 | Fischer's irises**  

##### **(a)** How many cases were included in the data?  

```{r}

data("iris")

nrow(iris)

```

There are 150 cases included in the data.
  

##### **(b)** How many numerical variables are included in the data? Indicate what they are, and if they are continuous or discrete.

```{r}

names(iris)

```

There are 4 numerical variables included in the dataset: Sepal Length, Sepal Width, Petal Length, and Petal Width. All of them are continuous.

  
##### **(c)** How many categorical variables are included in the data? Indicate what they are, and the corresponding levels (categories).

****

### **1.9 | Air pollution and birth outcomes, scope of inference** 

##### **(a)** Identify the population of interest and the sample in this study.

##### **(b)** Identify the population of interest and the sample in this study.

****

### **1.23 | Air pollution and birth outcomes, scope of inference** 


