---
title: "DATA 607, Week 3 -- Regular Expressions"
author: "Kavya Beheraj"
date: "February 17, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(stringr)
library(ggplot2)
```

***

The purpose of this assignment was to practice using regular expressions in R. These questions are from the end of Chapter 8 in the textbook *Automated Data Collection with R*.

***

### **3a:** Use the tools of this chapter to rearrange the vector so that all elements conform to the standard `first_name` `last_name`.

> **(i)** To start, I loaded the raw data and extracted the names.

```{r}

namesraw <- c("555-1239Moe Szyslak(636) 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert")

names <- unlist(str_extract_all(namesraw, "[[:alpha:]., ]{2,}"))

names

```

> **(ii)** Then, I removed the titles from the names, and trimmed the white space left over. I identified the titles and middle initial by looking for a pattern of 2 to 3 characters plus a period.

```{r}

trimmed <- str_replace_all(names, "[[:alpha:]]{2,3}\\.", "")

trimmed <- str_trim(trimmed)

trimmed

```

> **(iii)** To fix the names, my goal was to create clean lists of first names and last names that I could easily concatenate back together in the right order. I started by creating a clean list of last names. I identified last names as:
>
> * the pattern of all printable characters before a comma; OR
> * the specific pattern of 1 character + 1 blank space + any number of characters.
>
>Then, I removed the 1 character + 1 blank space pattern, and also removed the commas. This left me with a clean list of last names.

```{r}
last_names <- str_extract_all(trimmed,"[[:print:]]+[\\,]|[[:alpha:]][[:blank:]]{1}[[:alpha:]]+")

last_names <- str_replace_all(last_names,"[[:alpha:]]{1}[[:blank:]]{1}", "")

last_names <- str_replace_all(last_names, "\\,","")

last_names


```

> **(iv)** The clean list of first names was easier. I identified first names as:
>
> * the pattern of all printable characters after a comma, OR
> * the pattern of characters + blank spaces.
>
>Then, I removed the commas and trimmed the blank spaces, and got a clean list of first names.

```{r}
first_names <- str_extract_all(trimmed,"[\\,][[:print:]]+|[[:alpha:]]+[[:blank:]]+")

first_names <- str_replace_all(first_names,"[\\,]+[[:blank:]]{1,2}", "")

first_names <- str_trim(first_names)

first_names

```

> **(v)** Finally, I concatenated both lists in the right order and separated the names with a space.

```{r}
final <- str_c(first_names, last_names, sep=" ")

final

```


***

### **3b:** Construct a logical vector indicating whether the character has a title.

> A title is a pattern of two or three characters followed by a period. Two of the names fit this pattern.

```{r}

title <- grepl("[[:alpha:]]{2,3}\\.", names)

title

```

***

### **3c:** Construct a logical vector indicating whether the character has a second name.

> I used my cleaned list of first and last names to answer this question, since it only contains names (without titles). I identified a second name as characters followed by a period. Only one name fit this pattern.

```{r}

secondname <- grepl("[[:alpha:]]+\\.", final)

secondname

```

***

### Describe the types of strings that conform to the following regular expressions and construct an example that is matched by the regular expression.

### **4a:** `[0-9]+\\$`

> This pattern matches a string that has a number from 0 to 9 (at least once) followed by a dollar sign.
>
> Example: "4$"

```{r}

a1 <- c("4$", "4", "10$")

a2 <- grepl("[0-9]+\\$", a1)

a2

```

### **4b:** `\\b[a-z]{1,4}\\b`

> This pattern describes a string that has 1-4 letters from edge to edge.
>
> Example: "dddd"

```{r}

b1 <- c("hodc", "ff", "dddd", "5gs", "ggggg")

b2 <- grepl("\\b[a-z]{1,4}\\b", b1)

b2

```

### **4c:** `.*?\\.txt$`

> This pattern describes a string of any length (found at least once) that ends in ".txt".
>
> Example: "assignment.txt"

```{r}

c1 <- c("assignment.txt", "m.txt", ".txtm", "r.txt.y.txt")

c2 <- grepl(".*?\\.txt$", c1)

c2

```

### **4d:** `\\d{2}/\\d{2}/\\d{4}`

> This pattern describes a string with 2 digits, a forward slash, two digits, another forward slash, and 4 digits. It is constructed like a date.
>
> Example: "11/22/2018"

```{r}

d1 <- c("11222018", "11/22/2018")

d2 <- grepl("\\d{2}/\\d{2}/\\d{4}", d1)

d2

```

### **4e:** `<(.+?)>.+?</\\1>`

> This pattern matches an HTML tag with at least one character between the tags. It also has a backreference `\1`.
>
>Example: "`<b>bold</b>`"

```{r}

e1 <- c("<b>bold</b>", "<i>italics</i>")

e2 <- grepl("<(.+?)>.+?</\\1>", e1)

e2

```

***

### **9:** The following code hides a secret message. Crack it with R and regular expressions.

```{r}

secret <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0TanwoUwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigOd6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"

secret <- str_replace_all(secret, "[[:digit:]]", "")

secret <- str_replace_all(secret, "[[:lower:]]", "")

secret <- str_replace_all(secret, "\\.", " ")

secret

```

***
